---
title: "Soil Nutrient Exploratory Analysis"
author: "Abhishek Sinha"
date: "8/1/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2, quietly = T)
library(dplyr, quietly = T)
library(data.table, quietly = T)
```

##Data Loading

Loading data and converting strings to Factors
```{r}
system.time(soil.data <- fread("haryana_soil_nutrient.csv", header = T, stringsAsFactors = T))
```

##Basic Information

Structure of data
```{r}
str(soil.data)
```

Summary of data

```{r}
summary(soil.data)
```

##Missing Data

Find missing values in each column

```{r}
sapply(soil.data, function(x){sum(is.na(x))})
```

Find missing percentange of each column

```{r}
rows_tot <- nrow(soil.data)
sapply(soil.data, function(x){round(100*sum(is.na(x))/rows_tot, digits = 2)})

```

Boron, Nitrogen has more than 95% data missing, while Copper has 51% data missing. Problem is data for Nitrogen is missing.

NOTE: **Nitrogen** is one of the most important component in soil and missing data will create an issue for us.

###More details on Nitrogen

Block Wise
```{r}
nitrogen.missing <- is.na(soil.data$Nitrogen)
table(paste0(soil.data$DistrictName, " - ",soil.data$BlockName), nitrogen.missing)
```


District Wise

```{r}
table(soil.data$DistrictName, nitrogen.missing)
```

More Details

```{r}
nitrogen.data <- subset(soil.data, is.na(soil.data$Nitrogen)==FALSE)
nitrogen.data$Identity<-1
agg.district <-aggregate(nitrogen.data$Identity ~ nitrogen.data$DistrictName,data=nitrogen.data, FUN=sum)

qplot(agg.district$`nitrogen.data$DistrictName`, agg.district$`nitrogen.data$Identity`, data=agg.district, xlab = "District", ylab="Num of Entries with Nitrogen Data", col=`nitrogen.data$DistrictName`) + theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

This clearly shows that data is entered only in Rewari, Mewat and Faridabad.


##Electrical Conductivity

```{r}
ggplot(soil.data) + aes(x = DistrictName, y=ElectricalConductivity, fill=DistrictName) + geom_bar(stat = "summary", fun.y="mean")+ theme(axis.text.x = element_text(angle = 60, hjust = 1)) + xlab("District") + ylab("Electrical Conductivity")

```

High electrical conductivity of soil can be associated with more moisture in the soil. Palwal, Hisar, Faridabad and Jhajjar have high water content in the soil. 

Faridabad is close to Yamuna, so water content will be high. What about the other districts?  
* Good Irrigation  
* More Rainfall  

Need to find more details.
